---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
library(dplyr)
```

## Loading and preprocessing the data
The read below assumes that the working directory contains the data file called "activity.csv".

```{r readdata}
mydata <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
```{r stepsperday}
dailysteps <- aggregate(mydata$steps, by = list(mydata$date), FUN = sum)
colnames(dailysteps) <- c("date", "stepsum")
hist(dailysteps$stepsum,
     main="Histogram of Total Daily Steps Taken",
     xlab="Total Daily Steps",
     border="blue",
     col="green",
     las = 1,
     breaks = seq(0, 25000, 2500),
     ylim = c(0,20),
     xlim = c(0,25000))
meansteps <- mean(dailysteps$stepsum, na.rm = TRUE)
mediansteps <- median(dailysteps$stepsum, na.rm = TRUE)
```

The mean number of daily steps taken is **`r format(meansteps, scientific = FALSE)`** and the median is **`r mediansteps`**.

## What is the average daily activity pattern?
```{r averagedaily}
intervalmeans <- mydata %>%
  as_tibble() %>%
  group_by(interval) %>%
  summarise_at(vars(steps), funs(mean(., na.rm=TRUE)))
plot(intervalmeans$interval,
     intervalmeans$steps,
     type= "l",
     xlab= "5-minute interval",
     ylab= "Average number of steps taken",
     col= "green",
     lwd=2)
maxstepinterval <- intervalmeans$interval[which.max(intervalmeans$steps)]
```

Interval `r maxstepinterval` contains the maximum average number of steps taken.

## Imputing missing values
```{r missingvalues}
missingvaluesum <- sum(is.na(mydata))
#Replace all missing values with that day's average
imputeddata <- mydata
#imputeddata$steps[is.na(imputeddata$steps)] <- dailysteps$x[]
```

The total number of missing values in the dataset is `r missingvaluesum`.

## Are there differences in activity patterns between weekdays and weekends?
